---
title: "p8105_hw5_kpc2124"
author: "Kristi Chau"
date: "11/1/2019"
output: github_document
fig_caption: yes
latex_engine: xelatex
header-includes:
  - \usepackage{floatrow}
  - \floatsetup[figure]{capposition=top}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
library(knitr)
options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal())
```


# Problem 1 - iris dataset

```{r}
set.seed(10)

iris_with_missing = iris %>% 
  map_df(~replace(.x, sample(1:150, 20), NA)) %>%
  mutate(Species = as.character(Species))
```

Write a function that takes a vector as an argument; replaces missing values using the rules defined; and returns the resulting vector. Apply this function to the columns of iris_with_missing using a map statement.

Rules:
 - For numeric variables, you should fill in missing values with the mean of non-missing values
 - For character variables, you should fill in missing values with "virginica"

```{r}
iris_na = function(c) {
  if (is.numeric(c)) {
    replace_na(c, round(mean(c, na.rm = TRUE),2))
  }
  else if (is.character(c)) {
    replace_na(c, "virginica")
  }
}

output = 
  map(iris_with_missing,iris_na) %>% 
  bind_cols()

output
```


# Problem 2 - longitudinal study

This zip file contains data from a longitudinal study that included a control arm and an experimental arm. Data for each participant is included in a separate file, and file names include the subject ID and arm.

Create a tidy dataframe containing data from all participants, including the subject ID, arm, and observations over time:

Start with a dataframe containing all file names; the list.files function will help
Iterate over file names and read in data for each subject using purrr::map and saving the result as a new variable in the dataframe
Tidy the result; manipulate file names to include control arm and subject ID, make sure weekly observations are “tidy”, and do any other tidying that’s necessary
Make a spaghetti plot showing observations on each subject over time, and comment on differences between groups.







